# Combine collar data from Lotek and ATS collars

## Description
We use two brands of collars for my project, ATS and Lotek. Each brand of collar
has a slightly different spreadsheet output, so this code is needed to clean
and combine the data from ATS and Lotek collars.

This code reads in all the Lotek and ATS GPS .csv files, organizes them into
the same format, and combines them into one .csv with the associated cow
metadata (ear tag and breed). We need to do this before we create the database!

Jan 2023: Lotek GPS files are matched with cattle ID based only on collar ID.
This will be an issue in the future when we start re-using Lotek collars and
have multiple cattle IDs associated with each Lotek collar. When this happens,
I will need to add a section of code that looks not only at the collar ID but
also at the dates to match collars to cows.

```{r clean, include = F}
rm(list = ls())
```


```{r packages, include = F}
## Load necessary packages
library(dplyr)
library(data.table)
library(ggplot2)
library(purrr)
```

## Read Lotek collar data
```{r read_lotek}
# gps file names
lotek.collar.files.march =
  list.files(path = "../cattle/raw data/collar data/2022-03_collarData/gps",
             pattern = "*.csv", full.names = T)

lotek.collar.files.june =
  list.files(path = "../cattle/raw data/collar data/2022-06_collarData/gps", pattern = "*.csv", full.names = T)

lotek.collar.files.all = c(lotek.collar.files.march, lotek.collar.files.june)

# collar numbers
lotek.collar.numbers = lotek.collar.files.all %>%
  substring(nchar(.)-28, nchar(.)-24)

# collar data
lotek.collar.list = lotek.collar.files.all %>%
  map(~fread(.))

# add column with collar number
for(i in seq_along(lotek.collar.list)){
  lotek.collar.list[[i]]$CollarSerialNumber <- lotek.collar.numbers[i]
}

cattle.dat.lotek = lotek.collar.list %>%
  bind_rows()


```

## Read ATS data
```{r read_ats}
# gps file names
ats.files =
  list.files(path = "../cattle/raw data/collar data/ats", pattern = "*.txt", full.names = T)

# collar data
ats.collar.list = ats.files %>%
  map(~fread(.))

# create data.frame from list
cattle.dat.ATS = ats.collar.list %>%
  bind_rows()
```

## Combine collar data
```{r combine}
# what column names are we working with here?
names(cattle.dat.ATS)
names(cattle.dat.lotek)

### add date_time column to both data.frames ----
# Lotek
cattle.dat.lotek$date_time =
  as.POSIXct(cattle.dat.lotek$`GMT Time`,
             format = '%m/%d/%Y %H:%M:%S %p', tz = 'GMT')
# ATS
cattle.dat.ATS$Date = as.Date(cattle.dat.ATS$Julianday, 
                              origin = '2021-12-31')
cattle.dat.ATS$date_time = paste(cattle.dat.ATS$Hour, 
                                 cattle.dat.ATS$Minute, sep = ':') %>%
  paste(cattle.dat.ATS$Date, .) %>%
  as.POSIXct(format = '%Y-%m-%d %H:%M')

### rename NumSat column ----
cattle.dat.lotek = cattle.dat.lotek %>%
  rename(NumSats = Satellites)

### add DOP column to ATS
cattle.dat.ATS = cattle.dat.ATS %>%
  mutate(DOP = NA)

### add HDOP column to Lotek
cattle.dat.lotek = cattle.dat.lotek %>%
  mutate(HDOP = NA)

### drop unneeded columns ----
cattle.dat.ATS = cattle.dat.ATS %>%
  dplyr::select(c('CollarSerialNumber', 'date_time', 'Latitude', 'Longitude',
                  'Temperature', 'NumSats', 'DOP', 'HDOP'))
cattle.dat.lotek = cattle.dat.lotek %>%
  dplyr::select(c('CollarSerialNumber', 'date_time', 'Latitude', 'Longitude',
                  'Temperature', 'NumSats', 'DOP', 'HDOP'))

### create final data.frame ----
cattle.dat = rbind(cattle.dat.ATS, cattle.dat.lotek) %>%
  mutate(CollarSerialNumber = as.integer(CollarSerialNumber))

```

## add cow ear tag and breed info
```{r add_metadata}
## read in collar metadata
collar.metadata = 
  read.csv('../cattle/raw data/collar_pedometer-metadata/collar data log.csv') %>%
  mutate(download_start = as.POSIXct(download_start, format = '%m/%d/%Y')) %>%
  mutate(download_end = as.POSIXct(download_end, format = '%m/%d/%Y')) %>%
  rename(CollarSerialNumber = collar.no)

collar.metadataS22.S23 = collar.metadata %>%
  filter(download_end > as.POSIXct("2021-11-10 MST"))

cattle.dat.tags = left_join(cattle.dat, collar.metadataS22.S23[,1:3])

```

## clean collar data
NOTE: for now, I'm just removing all points with coordinates (0,0).
I will need to change this to filter points based on NumSats or DOP/HDOP
```{r clean_data}
cattle.dat.tags = cattle.dat.tags %>%
  filter(Latitude != 0 &
           Longitude != 0)
```

## save output to .csv
```{r save_csv, eval=F}
write.csv(cattle.dat.tags, file = '../processed data/2021-2022_GPS-combined.csv',
          row.names = F)
```
