# Create heat maps

## Description
I'm curious where the cows are hanging out, and if there is much difference 
between the areas inhabited by angus vs criollo cows. Before doing any advanced
analyses, my first step will be to plot the points from the cows' GPS collars,
overlaid on maps of the pastures in Indian Creek.

```{r clean3, include=FALSE}
rm(list = ls())
```

```{r packages3, include = F}
library(amt)
library(dplyr)
library(ggplot2)
library(leaflet)
library(lubridate)
library(mapview)
library(rgdal)
library(rlist)
library(sf)
```


## Load in pasture shapefiles
```{r shapefiles}
cottonwood = read_sf("./../spatial data/Cottonwood_Allotment_and_pasture_boundaries/Cottonwood_Allotment_and_pasture_boundaries.shp")

indianCreek = read_sf("./../spatial data/Cottonwood_Allotment_and_pasture_boundaries/BLM_IndianCreekAllotment.shp")

winterUSU = read_sf("./../spatial data/USU_study_winter_pastures_dissolve/USU_study_winter_pastures_dissolve.shp")
```

## Read in collar data, reproject to UTM, and convert to sf object
Here I'll read in the .csv file I created in Chapter 2 that contains all the GPS
collar data combined and organized.

```{r collar_dat}
# read in data
cattle.dat = read.csv('../cattle/processed data/2021-2022_GPS-combined.csv')

# convert to sf object
cattle.gps = st_as_sf(cattle.dat, coords = c('Longitude', 'Latitude'))
st_crs(cattle.gps) = CRS("+proj=longlat +ellps=WGS84")

# reproject to UTM
cattle.gps$geometry = st_transform(cattle.gps$geometry, 
                         crs = st_crs(cottonwood))

st_crs(cattle.gps) == st_crs(indianCreek)

# convert dates to posixct
cattle.gps$date_time = as.POSIXct(cattle.gps$date_time)

# convert breed to factor
cattle.gps$breed = as.factor(cattle.gps$breed)
```

## Calculate activity modes
Before I make the heatmaps, I'd like to classify the data by activity mode 
(namely, resting or grazing) based on the movement velocity recorded by the 
collars. That way, I can see if there there are differences within and between
breeds in the areas used for grazing or resting. I don't care quite so much
right now about the areas the cows just travel through, I really just want to
know where they're hanging out.

```{r activity_modes}
# create and clean cattle.frame
cattle.frame = (cattle.gps) %>%
  filter(!is.na(breed) &
           !is.na(date_time)) %>%
  mutate(tag.no = as.factor(tag.no)) %>%
  distinct()

# add x and y coordinates
cattle.frame =  bind_cols(cattle.frame,
                          as.data.frame(st_coordinates(cattle.frame)))

cattle.list = split(cattle.frame, f = cattle.frame$tag.no)
  
# make track
cattle.track = lapply(cattle.list, 
                      FUN = mk_track, .x = X, .y = Y, .t = date_time, 
                      crs = st_crs(cattle.gps), 
                      order_by_ts = T,
                      check_duplicates = F)
rm(cattle.list)

# calculate velocity from track
tags = levels(cattle.frame$tag.no)
cattle.speed = lapply(cattle.track, FUN = speed)
for(i in seq(1:length(tags))){
  c = tags[i]
  cattle.track[[c]][4] = cattle.speed[[c]]
  names(cattle.track[[c]])[4] = 'speed'
}

# convert list back into data.frame
speed.frame = list.rbind(cattle.track)
temp = array()
for(i in seq(1:length(tags))){
  temp = append(temp, rep(tags[i], times = dim(cattle.track[[i]])[1]))
}
temp = temp[!is.na(temp)]

speed.frame = speed.frame %>%
  mutate(tag.no = temp) %>%
  rename(X = x_, Y = y_, date_time = t_)

rm(cattle.track)

# merge cattle.frame and speed.frame
cattle.frame = merge(cattle.frame, speed.frame[,3:5], 
                       by = c('tag.no','date_time'))

# classify activity based on speed (from Nyamuryekung'e et al. 2021a)
resting = 2.34/60 # max resting speed in m/s
grazing = 25/60 # max grazing speed in m/s

cattle.frame = cattle.frame %>%
  mutate(activity = case_when(speed <= resting ~  'resting',
                            speed > resting & speed <= grazing ~
                              'grazing',
                            speed > grazing ~  'traveling',
                            T ~ 'none'))

cattle.grazing = cattle.frame %>%
  filter(activity == 'grazing')

cattle.resting = cattle.frame %>%
  filter(activity == 'resting')
```


```{r merge_sf, include=F}
### merge sf objects
cwood = cottonwood %>%
  dplyr::select(PASTURE_NA, geometry) %>%
  rename(Pasture = PASTURE_NA)

icreek = indianCreek %>%
  dplyr::select(Pasture, geometry)

USU = winterUSU %>%
  dplyr::select(USU_pastur, geometry) %>%
  rename(Pasture = USU_pastur)

pastures = rbind(cwood, USU)
```

## Create map of pastures
The cattle rotate through several pastures in Indian Creek from October through
May. This map shows roughly which months the cows were in each pasture in 2022.

```{r psature_schedule, include = F}
### add column for season in pastures sf
pastures = pastures %>%
  mutate(season = case_when(Pasture == 'Park Pasture' ~  'Jan-Feb',
                            Pasture == 'Creek Pasture' ~  'Mar',
                            Pasture == 'Drill Pasture' ~  'Apr',
                            Pasture == 'Davis' ~  'May',
                            Pasture == 'Lavender' ~  'May',
                            Pasture == 'Corral Pocket' ~ 'Nov-Dec',
                            T ~ 'Jun-Oct')) %>%
  mutate(season = ordered(season,
                             levels = c('Jan-Feb', 'Mar', 'Apr', 'May', 
                                        'Jun-Oct', 'Nov-Dec')))

### add shorthand pasture names
pastures = pastures %>%
  mutate(short.name = case_when(Pasture == 'Park Pasture' ~  'Park',
                            Pasture == 'Creek Pasture' ~  'Creek',
                            Pasture == 'Drill Pasture' ~  'Drill',
                            Pasture == 'Davis' ~  'Davis',
                            Pasture == 'Lavender' ~  'Lavender',
                            Pasture == 'Corral Pocket' ~ 'Corral Pocket',
                            T ~ 'NA')) %>%
  mutate(short.name = ordered(short.name,
                             levels = c('Park', 'Creek', 'Drill', 'Davis', 
                                        'Lavender', 'NA', 
                                        'Corral Pocket')))
### create data.frame with centroids
pastures = pastures %>%
  cbind(., st_coordinates(st_centroid(pastures)))

### create color palette for fill colors
colors = c('#99ddff', '#44bb99', '#bbcc33', '#aaaa00', '#eedd88', '#ee8866')

### make map
ggplot() + 
  geom_sf(data = pastures, aes(fill = season)) +
  geom_label(data = subset(pastures, short.name != 'NA'), 
                           aes(x = X, y = Y, label = short.name), size = 3.5) +
  scale_fill_manual(values = colors)

ggsave('../docs/yearly_rotation.png', plot = last_plot(), scale = 1)
```
<img src=`r markdown:::.b64EncodeFile("../docs/yearly_rotation.png")`>

## Interactive maps of each pasture
In this section, I have created Leaflet maps of each pasture. Feel free to zoom
in and out in each map!

### Park Pasture
```{r create_park_map, include = F}
# isolate pasture
park = subset(pastures, Pasture == 'Park Pasture')

# isolate points over pasture, transform to latlong (first rest, then graz)
park.points1 = st_intersection(park, cattle.resting, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

park.points2 = st_intersection(park, cattle.grazing, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

# transform park pasture
park = park %>% st_transform(CRS("+proj=longlat +datum=WGS84"))

factpal = colorFactor(c('#6c071c', '#00447c'), park.points1$breed)

# resting
park.map.resting = leaflet() %>%
  addPolygons(data = park, fill = NA, color = 'white') %>%
  addCircleMarkers(data = park.points1, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = park.points1, position = "topright",
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# grazing
park.map.grazing = leaflet() %>%
  addPolygons(data = park, fill = NA, color = 'white') %>%
  addCircleMarkers(data = park.points2, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = park.points2, position = "topright",
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")
```

```{r display_park_maps}
# Activity mode = resting
park.map.resting

# Activity mode = grazing
park.map.grazing

```

### Creek Pasture
```{r create_creek_map, include = F}
# isolate pasture
creek = subset(pastures, Pasture == 'Creek Pasture')

# isolate points over pasture, transform to latlong (first rest, then graz)
creek.points1 = st_intersection(creek, cattle.resting, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

creek.points2 = st_intersection(creek, cattle.grazing, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

# transform park pasture
creek = creek %>% st_transform(CRS("+proj=longlat +datum=WGS84"))

# resting
creek.map.resting = leaflet() %>%
  addPolygons(data = creek, fill = NA, color = 'white') %>%
  addCircleMarkers(data = creek.points1, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = creek.points1, position = "topright",
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# grazing
creek.map.grazing = leaflet() %>%
  addPolygons(data = creek, fill = NA, color = 'white') %>%
  addCircleMarkers(data = creek.points2, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = creek.points2, position = "topright",
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

```

```{r display_creek_maps}
# Activity mode = resting
creek.map.resting

# Activity mode = grazing
creek.map.grazing
```

### Drill Pasture
```{r create_drill_maps, include = F}
# isolate pasture
drill = subset(pastures, Pasture == 'Drill Pasture')

# isolate points over pasture, transform to latlong (first rest, then graz)
drill.points1 = st_intersection(drill, cattle.resting, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

drill.points2 = st_intersection(drill, cattle.grazing, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

# select every 12th point for plotting (10min -> 2hr interval)
drill.points1 = drill.points1[seq(1, nrow(drill.points1), 12),]
drill.points2 = drill.points1[seq(1, nrow(drill.points2), 12),]

# transform park pasture
drill = drill %>% st_transform(CRS("+proj=longlat +datum=WGS84"))

# resting
drill.map.resting = leaflet() %>%
  addPolygons(data = drill, fill = NA, color = 'white') %>%
  addCircleMarkers(data = drill.points1, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = drill.points1, position = "topright",
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# grazing
drill.map.grazing = leaflet() %>%
  addPolygons(data = drill, fill = NA, color = 'white') %>%
  addCircleMarkers(data = drill.points2, radius = .25,
                   color = ~factpal(breed), opacity = .5) %>%
  addLegend(data = drill.points2, position = "topright",
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

```

```{r display_drill_maps}
# Activity mode = resting
drill.map.resting

# Activity mode = grazing
drill.map.grazing
```

### Davis and Lavender Canyons
```{r create_davlav_maps, include = F}
# isolate pasture
davis.lav = subset(pastures, Pasture == 'Davis' |
                     Pasture == 'Lavender')

# isolate points over pasture, transform to latlong (first rest, then graz)
davis.lav.points1 = st_intersection(davis.lav, cattle.resting, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

davis.lav.points2 = st_intersection(davis.lav, cattle.grazing, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

# transform park pasture
davis.lav = davis.lav %>% st_transform(CRS("+proj=longlat +datum=WGS84"))

# resting
davis.lav.map.resting = leaflet() %>%
  addPolygons(data = davis.lav, fill = NA, color = 'white') %>%
  addCircleMarkers(data = davis.lav.points1, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = davis.lav.points1, position = "topright",
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# grazing
davis.lav.map.grazing = leaflet() %>%
  addPolygons(data = davis.lav, fill = NA, color = 'white') %>%
  addCircleMarkers(data = davis.lav.points2, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = davis.lav.points2, position = "topright",
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")
```

```{r display_davlav_maps}
# Activity mode = resting
davis.lav.map.resting

# Activity mode = grazing
davis.lav.map.grazing

```

### Corral Pocket
```{r create_corral_maps, include = F}
# isolate pasture
corral = subset(pastures, Pasture == 'Corral Pocket')

# isolate points over pasture, transform to latlong (first rest, then graz)
corral.points1 = st_intersection(corral, cattle.resting, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

corral.points2 = st_intersection(corral, cattle.grazing, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

# transform park pasture
corral = corral %>% st_transform(CRS("+proj=longlat +datum=WGS84"))

# resting
corral.map.resting = leaflet() %>%
  addPolygons(data = corral, fill = NA, color = 'white') %>%
  addCircleMarkers(data = corral.points1, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = corral.points1, position = "topright",
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# grazing
corral.map.grazing = leaflet() %>%
  addPolygons(data = corral, fill = NA, color = 'white') %>%
  addCircleMarkers(data = corral.points2, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = corral.points2, position = "topright",
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

```

```{r display_corral_maps}
# Activity mode = resting
corral.map.resting

# Activity mode = grazing
corral.map.grazing

```
