---
title: "full year heat map 2022"
author: "Maria Stahl"
date: "1/19/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
```

### load packages
```{r}
library(amt)
library(dplyr)
library(ggplot2)
library(leaflet)
library(lubridate)
library(mapview)
library(rgdal)
library(rlist)
library(sf)
```


## Load in pasture shapefiles
```{r}
cottonwood = read_sf("./../../spatial data/Cottonwood_Allotment_and_pasture_boundaries/Cottonwood_Allotment_and_pasture_boundaries.shp")

indianCreek = read_sf("./../../spatial data/Cottonwood_Allotment_and_pasture_boundaries/BLM_IndianCreekAllotment.shp")

winterUSU = read_sf("./../../spatial data/USU_study_winter_pastures_dissolve/USU_study_winter_pastures_dissolve.shp")
```

### Read in collar data, reproject to UTM, and convert to sf object
```{r}
# read in data
cattle.dat = read.csv('../processed data/2021-2022_GPS-combined.csv')

# convert to sf object
cattle.gps = st_as_sf(cattle.dat, coords = c('Longitude', 'Latitude'))
st_crs(cattle.gps) = CRS("+proj=longlat +ellps=WGS84")

# reproject to UTM
cattle.gps$geometry = st_transform(cattle.gps$geometry, 
                         crs = st_crs(cottonwood))

st_crs(cattle.gps) == st_crs(indianCreek)

# convert dates to posixct
cattle.gps$date_time = as.POSIXct(cattle.gps$date_time)

# convert breed to factor
cattle.gps$breed = as.factor(cattle.gps$breed)
```

## Calculate activity modes
```{r}
# create and clean cattle.frame
cattle.frame = (cattle.gps) %>%
  filter(!is.na(breed) &
           !is.na(date_time)) %>%
  mutate(tag.no = as.factor(tag.no)) %>%
  distinct()

# add x and y coordinates
cattle.frame =  bind_cols(cattle.frame,
                          as.data.frame(st_coordinates(cattle.frame)))

cattle.list = split(cattle.frame, f = cattle.frame$tag.no)
  
# make track
cattle.track = lapply(cattle.list, 
                      FUN = mk_track, .x = X, .y = Y, .t = date_time, 
                      crs = st_crs(cattle.gps), 
                      order_by_ts = T,
                      check_duplicates = F)
rm(cattle.list)

# calculate velocity from track
tags = levels(cattle.frame$tag.no)
cattle.speed = lapply(cattle.track, FUN = speed)
for(i in seq(1:length(tags))){
  c = tags[i]
  cattle.track[[c]][4] = cattle.speed[[c]]
  names(cattle.track[[c]])[4] = 'speed'
}

# convert list back into data.frame
speed.frame = list.rbind(cattle.track)
temp = array()
for(i in seq(1:length(tags))){
  temp = append(temp, rep(tags[i], times = dim(cattle.track[[i]])[1]))
}
temp = temp[!is.na(temp)]

speed.frame = speed.frame %>%
  mutate(tag.no = temp) %>%
  rename(X = x_, Y = y_, date_time = t_)

rm(cattle.track)

# merge cattle.frame and speed.frame
cattle.frame = merge(cattle.frame, speed.frame[,3:5], 
                       by = c('tag.no','date_time'))

# classify activity based on speed (from Nyamuryekung'e et al. 2021a)
resting = 2.34/60 # max resting speed in m/s
grazing = 25/60 # max grazing speed in m/s

cattle.frame = cattle.frame %>%
  mutate(activity = case_when(speed <= resting ~  'resting',
                            speed > resting & speed <= grazing ~
                              'grazing',
                            speed > grazing ~  'traveling',
                            T ~ 'none'))

cattle.grazing = cattle.frame %>%
  filter(activity == 'grazing')

cattle.resting = cattle.frame %>%
  filter(activity == 'resting')
```

### plot all points on top of pastures
```{r}
# ggplot() +
#   geom_sf(data = cottonwood) +
#   geom_sf(data = indianCreek) +
#   geom_sf(data = cattle.gps, alpha = 0.5)

```

### merge sf objects
```{r}
cwood = cottonwood %>%
  dplyr::select(PASTURE_NA, geometry) %>%
  rename(Pasture = PASTURE_NA)

icreek = indianCreek %>%
  dplyr::select(Pasture, geometry)

USU = winterUSU %>%
  dplyr::select(USU_pastur, geometry) %>%
  rename(Pasture = USU_pastur)

pastures = rbind(cwood, USU)
```

### remove points outside of pasture boundaries
```{r}
# points.in = st_intersection(pastures, cattle.gps, sparse = F)
```

### make smaller cattle points data.frame for faster plotting
```{r}
# # select every 100th point from points.in
# points.in.small = points.in[seq(1, nrow(points.in), 100),]
```

### plot small points df on top of pastures
```{r}

# # add column for 'month' in points.in.small
# points.in.small = points.in.small %>% 
#   mutate(month = month(date_time))
# 
# ggplot() +
#   geom_sf(data = pastures, fill = 'white', linewidth = 1.25) +
#   # geom_sf(data = subset(points.in.small),
#   #         aes(color = as.factor(month)), alpha = 0.5) +
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#         panel.background = element_blank(), 
#         axis.line = element_line(colour = "black"))

```

### create color-coded map of pastures
```{r}
### add column for season in pastures sf
pastures = pastures %>%
  mutate(season = case_when(Pasture == 'Park Pasture' ~  'Jan-Feb',
                            Pasture == 'Creek Pasture' ~  'Mar',
                            Pasture == 'Drill Pasture' ~  'Apr',
                            Pasture == 'Davis' ~  'May',
                            Pasture == 'Lavender' ~  'May',
                            Pasture == 'Corral Pocket' ~ 'Nov-Dec',
                            T ~ 'Jun-Oct')) %>%
  mutate(season = ordered(season,
                             levels = c('Jan-Feb', 'Mar', 'Apr', 'May', 
                                        'Jun-Oct', 'Nov-Dec')))

### add shorthand pasture names
pastures = pastures %>%
  mutate(short.name = case_when(Pasture == 'Park Pasture' ~  'Park',
                            Pasture == 'Creek Pasture' ~  'Creek',
                            Pasture == 'Drill Pasture' ~  'Drill',
                            Pasture == 'Davis' ~  'Davis',
                            Pasture == 'Lavender' ~  'Lavender',
                            Pasture == 'Corral Pocket' ~ 'Corral Pocket',
                            T ~ 'NA')) %>%
  mutate(short.name = ordered(short.name,
                             levels = c('Park', 'Creek', 'Drill', 'Davis', 
                                        'Lavender', 'NA', 
                                        'Corral Pocket')))
### create data.frame with centroids
pastures = pastures %>%
  cbind(., st_coordinates(st_centroid(pastures)))

### create color palette for fill colors
colors = c('#99ddff', '#44bb99', '#bbcc33', '#aaaa00', '#eedd88', '#ee8866')

### make map
ggplot() + 
  geom_sf(data = pastures, aes(fill = season)) +
  geom_label(data = subset(pastures, short.name != 'NA'), 
                           aes(x = X, y = Y, label = short.name), size = 3.5) +
  scale_fill_manual(values = colors)

# ggsave('../figures/yearly rotation.png', plot = last_plot(), scale = 1)
```

### look at each pasture separately for spatial partitioning
## PARK
```{r}
# isolate pasture
park = subset(pastures, Pasture == 'Park Pasture')

# isolate points over pasture, transform to latlong (first rest, then graz)
park.points1 = st_intersection(park, cattle.resting, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

park.points2 = st_intersection(park, cattle.grazing, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

# transform park pasture
park = park %>% st_transform(CRS("+proj=longlat +datum=WGS84"))

factpal = colorFactor(c('#6c071c', '#00447c'), park.points1$breed)

# resting
park.map.resting = leaflet() %>% 
  addPolygons(data = park, fill = NA, color = 'white') %>%
  addCircleMarkers(data = park.points1, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = park.points1, position = "topright", 
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# mapshot(park.map.resting, 
#         url = '../figures/2021-2022 leaflet maps/parkMapResting.html',
#         file = '../figures/2021-2022 leaflet maps/parkMapResting.png')

# grazing
park.map.grazing = leaflet() %>% 
  addPolygons(data = park, fill = NA, color = 'white') %>%
  addCircleMarkers(data = park.points2, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = park.points2, position = "topright", 
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")
# 
# mapshot(park.map.grazing, 
#         url = '../figures/2021-2022 leaflet maps/parkMapGrazing.html',
#         file = '../figures/2021-2022 leaflet maps/parkMapGrazing.png')

```

## CREEK
```{r}
# isolate pasture
creek = subset(pastures, Pasture == 'Creek Pasture')

# isolate points over pasture, transform to latlong (first rest, then graz)
creek.points1 = st_intersection(creek, cattle.resting, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

creek.points2 = st_intersection(creek, cattle.grazing, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

# transform park pasture
creek = creek %>% st_transform(CRS("+proj=longlat +datum=WGS84"))

# resting
creek.map.resting = leaflet() %>% 
  addPolygons(data = creek, fill = NA, color = 'white') %>%
  addCircleMarkers(data = creek.points1, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = creek.points1, position = "topright", 
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# mapshot(creek.map.resting, 
#         url = '../figures/2021-2022 leaflet maps/creekMapResting.html',
#         file = '../figures/2021-2022 leaflet maps/creekMapResting.png')

# grazing
creek.map.grazing = leaflet() %>% 
  addPolygons(data = creek, fill = NA, color = 'white') %>%
  addCircleMarkers(data = creek.points2, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = creek.points2, position = "topright", 
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# mapshot(creek.map.grazing, 
#         url = '../figures/2021-2022 leaflet maps/creekMapGrazing.html',
#         file = '../figures/2021-2022 leaflet maps/creekMapGrazing.png')

```

## DRILL
```{r}
# isolate pasture
drill = subset(pastures, Pasture == 'Drill Pasture')

# isolate points over pasture, transform to latlong (first rest, then graz)
drill.points1 = st_intersection(drill, cattle.resting, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

drill.points2 = st_intersection(drill, cattle.grazing, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

# select every 12th point for plotting (10min -> 2hr interval)
drill.points1 = drill.points1[seq(1, nrow(drill.points1), 12),]
drill.points2 = drill.points1[seq(1, nrow(drill.points2), 12),]

# transform park pasture
drill = drill %>% st_transform(CRS("+proj=longlat +datum=WGS84"))

# resting
drill.map.resting = leaflet() %>% 
  addPolygons(data = drill, fill = NA, color = 'white') %>%
  addCircleMarkers(data = drill.points1, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = drill.points1, position = "topright", 
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# mapshot(drill.map.resting, 
#         url = '../figures/2021-2022 leaflet maps/drillMapResting.html',
#         file = '../figures/2021-2022 leaflet maps/drillMapResting.png')

# grazing
drill.map.grazing = leaflet() %>% 
  addPolygons(data = drill, fill = NA, color = 'white') %>%
  addCircleMarkers(data = drill.points2, radius = .25,
                   color = ~factpal(breed), opacity = .5) %>%
  addLegend(data = drill.points2, position = "topright", 
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# mapshot(drill.map.grazing, 
#         url = '../figures/2021-2022 leaflet maps/drillMapGrazing.html',
#         file = '../figures/2021-2022 leaflet maps/drillMapGrazing.png')

```

## DAVIS AND LAVENDER
```{r}
# isolate pasture
davis.lav = subset(pastures, Pasture == 'Davis' |
                     Pasture == 'Lavender')

# isolate points over pasture, transform to latlong (first rest, then graz)
davis.lav.points1 = st_intersection(davis.lav, cattle.resting, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

davis.lav.points2 = st_intersection(davis.lav, cattle.grazing, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

# transform park pasture
davis.lav = davis.lav %>% st_transform(CRS("+proj=longlat +datum=WGS84"))

# resting
davis.lav.map.resting = leaflet() %>% 
  addPolygons(data = davis.lav, fill = NA, color = 'white') %>%
  addCircleMarkers(data = davis.lav.points1, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = davis.lav.points1, position = "topright", 
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# mapshot(davis.lav.map.resting, 
#         url = '../figures/2021-2022 leaflet maps/davis.lavMapResting.html',
#         file = '../figures/2021-2022 leaflet maps/davis.lavMapResting.png')

# grazing
davis.lav.map.grazing = leaflet() %>% 
  addPolygons(data = davis.lav, fill = NA, color = 'white') %>%
  addCircleMarkers(data = davis.lav.points2, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = davis.lav.points2, position = "topright", 
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# mapshot(davis.lav.map.grazing, 
#         url = '../figures/2021-2022 leaflet maps/davis.lavMapGrazing.html',
#         file = '../figures/2021-2022 leaflet maps/davis.lavMapGrazing.png')

```

## CORRAL POCKET
```{r}
# isolate pasture
corral = subset(pastures, Pasture == 'Corral Pocket')

# isolate points over pasture, transform to latlong (first rest, then graz)
corral.points1 = st_intersection(corral, cattle.resting, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

corral.points2 = st_intersection(corral, cattle.grazing, sparse = F) %>%
  st_transform(CRS("+proj=longlat +datum=WGS84"))

# transform park pasture
corral = corral %>% st_transform(CRS("+proj=longlat +datum=WGS84"))

# resting
corral.map.resting = leaflet() %>% 
  addPolygons(data = corral, fill = NA, color = 'white') %>%
  addCircleMarkers(data = corral.points1, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = corral.points1, position = "topright", 
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# mapshot(corral.map.resting, 
#         url = '../figures/2021-2022 leaflet maps/corralMapResting.html',
#         file = '../figures/2021-2022 leaflet maps/corralMapResting.png')

# grazing
corral.map.grazing = leaflet() %>% 
  addPolygons(data = corral, fill = NA, color = 'white') %>%
  addCircleMarkers(data = corral.points2, radius = .25,
                   color = ~factpal(breed), opacity = .25) %>%
  addLegend(data = corral.points2, position = "topright", 
            pal = factpal, values = ~breed,
            opacity = 1) %>%
  addProviderTiles("Esri.WorldImagery")

# mapshot(corral.map.grazing, 
#         url = '../figures/2021-2022 leaflet maps/corralMapGrazing.html',
#         file = '../figures/2021-2022 leaflet maps/corralMapGrazing.png')
#        selfcontained = FALSE)

```

